# Assignmnet 1
===============================================================================================================
---
title: "Assignmnet 1"
author: "Alan"
date: "January 24, 2016"
output: html_document
---

This is the R Markdown document for Peer Assignment 1, Reproducible Research Course.
```{r setoptions, echo=TRUE}
```
This was the first project for the Exploratory Data Analysis course in Coursera's Data Science specialization track. The objective was to practice creating plots by reproducing a series of four plots, and also to practice using github by forking and cloning an existing repository.
###Data Loading and Preprocessing
```{r}

power <- read.table("household_power_consumption.txt", header=T, sep=";")
finalData <- power[power$Date %in% c("1/2/2007","2/2/2007"),]
SetTime <-strptime(paste(finalData$Date, finalData$Time, sep=" "),"%d/%m/%Y %H:%M:%S")
finalData <- cbind(SetTime, finalData)
finalData$Global_active_power <- as.numeric(as.character(finalData$Global_active_power))
finalData$Sub_metering_1 <- as.numeric(as.character(finalData$Sub_metering_1))
finalData$Sub_metering_2 <- as.numeric(as.character(finalData$Sub_metering_2))
finalData$Sub_metering_3 <- as.numeric(as.character(finalData$Sub_metering_3))
```
### Generating Plot 1
```{r}
plot1 <- function(){
hist(finalData$Global_active_power, col="red", main="Global Active Power", xlab="Global Active Power (kilowatts)")
dev.copy(png, file="plot1.png", width=480, height=480)
dev.off()
 cat("Plot1.png has been saved in", getwd())
}
plot1()

```
### Generating Plot 2
```{r}
plot2 <- function() {
plot(finalData$SetTime, finalData$Global_active_power, type="l", col="black", xlab="", ylab="Global Active Power (kilowatts)")
  dev.copy(png, file="plot2.png", width=480, height=480)
  dev.off()
  cat("plot2.png has been saved in", getwd())
}
plot2()

```
### Generating Plot 3
```{r}
plot3 <- function() {
columnlines <- c("black", "red", "blue")
labels <- c("Sub_metering_1", "Sub_metering_2", "Sub_metering_3")
plot(finalData$SetTime, finalData$Sub_metering_1, type="l", col=columnlines[1], xlab="", ylab="Energy sub metering")
lines(finalData$SetTime, finalData$Sub_metering_2, col=columnlines[2])
lines(finalData$SetTime, finalData$Sub_metering_3, col=columnlines[3])
legend("topright", legend=labels, col=columnlines, lty="solid")
dev.copy(png, file="plot3.png", width=480, height=480)
dev.off()
cat("plot3.png has been saved in", getwd())
}
plot3()

```
### Generating Plot 4
```{r}
plot4 <- function() {
  par(mfrow=c(2,2))
  
  ## subPLOT 1
  plot(finalData$SetTime,finalData$Global_active_power, type="l", xlab="", ylab="Global Active Power")
  ## subPLOT 2
  plot(finalData$SetTime,finalData$Voltage, type="l", xlab="datetime", ylab="Voltage")
  
  ## subPLOT 3
  plot(finalData$SetTime,finalData$Sub_metering_1, type="l", xlab="", ylab="Energy sub metering")
  lines(finalData$SetTime,finalData$Sub_metering_2,col="red")
  lines(finalData$SetTime,finalData$Sub_metering_3,col="blue")
  legend("topright", col=c("black","red","blue"), c("Sub_metering_1  ","Sub_metering_2  ", "Sub_metering_3  "),lty=c(1,1), bty="n", cex=.5) #bty removes the box, cex shrinks the text, spacing added after labels so it renders correctly
  
  # subPLOT 4
  plot(finalData$SetTime,finalData$Global_reactive_power, type="l", xlab="datetime", ylab="Global_reactive_power")
  
  #OUTPUT
  dev.copy(png, file="plot4.png", width=480, height=480)
  dev.off()
  cat("plot4.png has been saved in", getwd())
}
plot4()
```
